Addressing the reviser concerns

### 2019-09-27 draft/preprint_JCAP_v0.2/paper.tex
implemented more changes from revision to v0.1

      The pseudo-$C_\ell$ estimator then proceeds in two steps:
      \begin{enumerate}
        \item We first bin different ${\bf l}$ modes into sets of them called bandpowers (typically bands of similar $\ell$ or annuli of flat-sky Fourier modes spanning a range of radii). Let us denote a given bandpower by its index $q$. The binned pseudo-power spectrum is:
        \begin{equation}\label{eq:bandpowers}
          \tilde{\sf C}^{ab}_q=\sum_{{\bf l}\in q} B_q^{\bf l}\,\tilde{\bf a}_{\bf l}\tilde{\bf b}^\dag_{\bf l},
        \end{equation}
        where the bandpower weights are normalized such that $\sum_{{\bf l}\in
          q}B_q^{\bf l}=(K\Delta^x({\bf 0}))^{-1}$.
        \item Then, the correlation between bandpowers induced by the mode-coupling coefficients is partially reversed by multiplying $\tilde{\sf C}^{ab}_q$ by the so-called binned ``mode-coupling matrix'' $\mathcal{M}$, giving the final estimator
        \begin{equation}
          {\rm vec}\left(\hat{\sf C}^{ab}_q\right) = \sum_{q'} \left(\mathcal{M}^{-1}\right)_{qq'}\,{\rm vec}\left(\tilde{\sf C}^{ab}_{q'}\right).
        \end{equation}

@@

      For our fiducial results, we generated a set of 20,000 random
      simulations. This number was chosen in order to recover the covariance
      matrix for all possible auto- and cross-correlations with sufficient
      accuracy. All simulations were generated as {\tt
        HEALPix}\footnote{\url{http://healpix.sourceforge.net}} maps with resolution $N_{\rm side}=512$. We then used {\tt NaMaster} to compute all possible power spectra for each simulation using narrow bandpowers of width $\Delta\ell=3$ from $\ell=2$ to $\ell=1023$. Finally, we used the power spectra from the simulations to estimate the sample covariance matrix
      \begin{equation}
        {\rm Cov}({\bf C}) =\frac{1}{N_{\rm sim}-1}\sum_{i=1}^{N_{\rm sim}}\left({\bf C}_i-\bar{\bf C}\right)\cdot\left({\bf C}_i-\bar{\bf C}\right)^T,
      \end{equation}
      where ${\bf C}_i$ is the vector of all possible power spectra for the $i$-th realization, and $\bar{\bf C}$ is the mean of this vector over all realizations. The comparison of this sample covariance with the analytical approximations described above is presented in the next sections.

    \subsection{Qualitative comparison}\label{ssec:results.visual}
      \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{./figures/all_rows_sph_1bin.pdf}
        \caption{Four rows of the covariance matrix for different power spectra. The top panels show the cases of spectra with non-zero signal, involving $\delta$ and $\gamma_E$, while the bottom panels show cases involving $B$-modes. In each panel we show rows of the covariance matrix for $\ell=60,\,90,\,120$ and $150$, which peak at those central values. The different lines show the results for the sample covariance matrix (solid blue), its NKA estimator (dashed orange) and the difference between both (dotted green). The black stars show the mode-counting approximation to the covariance matrix (Eq. \ref{eq:cov_naive_bpw}). To facilitate the visualization of the different rows, we have divided them by $C_\ell^{aa}C_\ell^{bb}+(C^{ab}_\ell)^2$, the numerator of Eq.~\ref{eq:cov_naive_bpw}, for the two fields, $a$ and $b$, that are being correlated in each case. The NKA method is able to recover the covariance with high accuracy for all field combinations that do not involve $B$ modes, yielding visibly poorer results otherwise.}
        \label{fig:rows_1bin}
      \end{figure}
      \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{./figures/sph_1bin_diff_corr.pdf}
        \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\boldsymbol{\gamma}$ in a single redshift bin. While the NKA estimator is able to recover the covariance matrix to high accuracy in most cases, it is not able to reproduce the off-diagonal correlations between different bandpowers in cases involving $B$ modes.} \label{fig:diff_corr_1bin}
      \end{figure}
      \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{./figures/run_sph_2b_NKA_diff_corr_short.pdf}
        \caption{Same as Fig. \ref{fig:diff_corr_1bin} for all cross-correlations between $\delta$ and $\gamma_E$ measured in two different redshift bins.}
        \label{fig:corr_diff_2bins}
      \end{figure}
      As a first step, we visually compare the main properties of the sample covariance matrix estimated from the simulations and the NKA, Spin-0 and MC approximations described in the previous section. Figure \ref{fig:rows_1bin} shows four rows of the covariance matrix of different auto- and cross-correlations. The rows correspond to bandpowers centered on multipoles $\ell_q\simeq60,\,90,\,120$ and $150$. The upper panels show results for the non-zero power spectra ($\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$), with the solid blue, dashed orange and dotted green lines showing results for the sample covariance matrix, the analytical covariance using the NKA approximation and their difference, respectively. For comparison, the black stars show the diagonal covariance matrix elements predicted by the MC approximation (Eq. \ref{eq:cov_naive_bpw}). We find an excellent agreement between the simulated and analytical covariances, with very small deviations in the amplitude of the diagonal and first few off-diagonal elements.

@@

    \subsection{Impact on parameter estimation}\label{ssec:results.cosmo}
      \begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{./figures/contours_2bin.pdf}
        \caption{1$\sigma$ and 2$\sigma$ contours for $\Omega_M$ and $\sigma_8$ found from two different estimates of the power spectrum likelihood for a dataset containing all possible cross-power spectra between $\delta$ and $\gamma_E$ for two redshift bins. The filled contours correspond to the parameter likelihood evaluated using the NKA estimator for the power spectrum covariance matrix, while the dashed black contours correspond to the same calculation done with the sample covariance matrix corrected by the factor in Eq. \ref{eq:hartlap}. In both cases we assume a Gaussian likelihood. We find a remarkably good agreement between both likelihoods, highlighting the negligible impact of the approximations involved in the NKA method on the final parameter estimates.} \label{fig:contours}
      \end{figure}
      Ultimately, the most important test to judge the accuracy of the
      analytical covariance matrix estimators implemented here is to study
      their impact on the posterior distribution of cosmological parameters
      derived from power spectrum measurements. Assuming flat priors \cgg{and
      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}}, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
      \begin{equation}\label{eq:parpos}
        -2\log p(\vec{\theta}|\hat{\bf C})=\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)^T\cdot{\rm Cov}^{-1}\cdot\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)+{\rm const.},
      \end{equation}
      where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.

@@

  \section{Discussion}\label{sec:discussion}
    Estimating accurate covariance matrices for projected two-point correlators is an ubiquitous problem in modern cosmology \cite{2013PhRvD..88f3537D,2013MNRAS.432.1928T,2017arXiv170609359K}, but it is particularly relevant for large-scale structure datasets. This problem is further complicated in this case, in comparison with e.g. CMB experiments, by two factors: the fact that the fields involved are non-Gaussian at some level and the higher complexity of the sky masks used in optical datasets. The main impact of survey geometry is the statistical coupling it induces between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable estimates of the posterior distribution for cosmological parameters. In this paper we have focused on the impact of survey geometry on the dominant Gaussian (i.e. disconnected) part of the covariance matrix.

    We have described and generalized existing analytical approaches to estimate the covariance matrix for pseudo-$C_\ell$ power spectrum estimators \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and implemented them in the public code {\tt NaMaster} \cite{2019MNRAS.484.4127A}, making it straightforward to fully account for the effects of survey geometry on the data uncertainties. With these approximation, computationally speaking, the problem of estimating a covariance matrix is as complex as that of computing the power spectrum itself, and scales with the number of pixels in the map as $N_{\rm pix}^{3/2}$.

    The main finding of this paper is the excellent performance of the analytical methods described in Section \ref{ssec:theory.pclcov}. We have shown that, in spite of the higher complexity of large-scale structure masks, the NKA estimator (see Section \ref{ssec:theory.approx}) is able to recover the covariance matrix for all power spectra with non-zero signal expectation value (i.e. those involving the galaxy overdensity $\delta$ or the $E$-mode shear $\gamma_E$), as well as the posterior distribution for cosmological parameters, to a high degree of accuracy.

>>>>

      The pseudo-$C_\ell$ estimator then proceeds in two steps:
      \begin{enumerate}
        \item We first bin different ${\bf l}$ modes into sets of them called
          bandpowers (typically bands of similar $\ell$ or annuli of flat-sky
          Fourier modes spanning a range of radii). Let us denote a given
          bandpower by its index $q$. \cgg{The number of bandpowers and their
            width are determined by the true underlying power spectra, which
            must not vary much within the bandpower}. The binned pseudo-power spectrum is:
        \begin{equation}\label{eq:bandpowers}
          \tilde{\sf C}^{ab}_q=\sum_{{\bf l}\in q} B_q^{\bf l}\,\tilde{\bf a}_{\bf l}\tilde{\bf b}^\dag_{\bf l},
        \end{equation}
        where the bandpower weights are normalized such that $\sum_{{\bf l}\in
          q}B_q^{\bf l}=(K\Delta^x({\bf 0}))^{-1}$.
        \item Then, the correlation between bandpowers induced by the mode-coupling coefficients is partially reversed by multiplying $\tilde{\sf C}^{ab}_q$ by the so-called binned ``mode-coupling matrix'' $\mathcal{M}$, giving the final estimator
        \begin{equation}
          {\rm vec}\left(\hat{\sf C}^{ab}_q\right) = \sum_{q'} \left(\mathcal{M}^{-1}\right)_{qq'}\,{\rm vec}\left(\tilde{\sf C}^{ab}_{q'}\right).
        \end{equation}

@@

      For our fiducial results, we generated a set of 20,000 random
      simulations. This number was chosen in order to recover the covariance
      matrix for all possible auto- and cross-correlations with sufficient
      accuracy. All simulations were generated as {\tt
        HEALPix}\footnote{\url{http://healpix.sourceforge.net}}~\cite{Gorski:2004by} maps with resolution $N_{\rm side}=512$. We then used {\tt NaMaster} to compute all possible power spectra for each simulation using narrow bandpowers of width $\Delta\ell=3$ from $\ell=2$ to $\ell=1023$. Finally, we used the power spectra from the simulations to estimate the sample covariance matrix
      \begin{equation}
        {\rm Cov}({\bf C}) =\frac{1}{N_{\rm sim}-1}\sum_{i=1}^{N_{\rm sim}}\left({\bf C}_i-\bar{\bf C}\right)\cdot\left({\bf C}_i-\bar{\bf C}\right)^T,
      \end{equation}
      where ${\bf C}_i$ is the vector of all possible power spectra for the $i$-th realization, and $\bar{\bf C}$ is the mean of this vector over all realizations. The comparison of this sample covariance with the analytical approximations described above is presented in the next sections.

    \subsection{Qualitative comparison}\label{ssec:results.visual}
      \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{./figures/all_rows_sph_1bin.pdf}
        \caption{Four rows of the covariance matrix for different power spectra. The top panels show the cases of spectra with non-zero signal, involving $\delta$ and $\gamma_E$, while the bottom panels show cases involving $B$-modes. In each panel we show rows of the covariance matrix for $\ell=60,\,90,\,120$ and $150$, which peak at those central values. The different lines show the results for the sample covariance matrix (solid blue), its NKA estimator (dashed orange) and the difference between both (dotted green). The black stars show the mode-counting approximation to the covariance matrix (Eq. \ref{eq:cov_naive_bpw}). To facilitate the visualization of the different rows, we have divided them by $C_\ell^{aa}C_\ell^{bb}+(C^{ab}_\ell)^2$, the numerator of Eq.~\ref{eq:cov_naive_bpw}, for the two fields, $a$ and $b$, that are being correlated in each case. The NKA method is able to recover the covariance with high accuracy for all field combinations that do not involve $B$ modes, yielding visibly poorer results otherwise.}
        \label{fig:rows_1bin}
      \end{figure}
      \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{./figures/sph_1bin_diff_corr-lims-v02.pdf}
        \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\boldsymbol{\gamma}$ in a single redshift bin. While the NKA estimator is able to recover the covariance matrix to high accuracy in most cases, it is not able to reproduce the off-diagonal correlations between different bandpowers in cases involving $B$ modes.} \label{fig:diff_corr_1bin}
      \end{figure}
      \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{./figures/run_sph_2b_NKA_diff_corr_short-lims-v02.pdf}
        \caption{Same as Fig. \ref{fig:diff_corr_1bin} for all cross-correlations between $\delta$ and $\gamma_E$ measured in two different redshift bins.}
        \label{fig:corr_diff_2bins}
      \end{figure}
      As a first step, we visually compare the main properties of the sample covariance matrix estimated from the simulations and the NKA, Spin-0 and MC approximations described in the previous section. Figure \ref{fig:rows_1bin} shows four rows of the covariance matrix of different auto- and cross-correlations. The rows correspond to bandpowers centered on multipoles $\ell_q\simeq60,\,90,\,120$ and $150$. The upper panels show results for the non-zero power spectra ($\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$), with the solid blue, dashed orange and dotted green lines showing results for the sample covariance matrix, the analytical covariance using the NKA approximation and their difference, respectively. For comparison, the black stars show the diagonal covariance matrix elements predicted by the MC approximation (Eq. \ref{eq:cov_naive_bpw}). We find an excellent agreement between the simulated and analytical covariances, with very small deviations in the amplitude of the diagonal and first few off-diagonal elements.

@@

    \subsection{Impact on parameter estimation}\label{ssec:results.cosmo}
      \begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{./figures/contours_2bin.pdf}
        \caption{1$\sigma$ and 2$\sigma$ contours for $\Omega_M$ and $\sigma_8$ found from two different estimates of the power spectrum likelihood for a dataset containing all possible cross-power spectra between $\delta$ and $\gamma_E$ for two redshift bins. The filled contours correspond to the parameter likelihood evaluated using the NKA estimator for the power spectrum covariance matrix, while the dashed black contours correspond to the same calculation done with the sample covariance matrix corrected by the factor in Eq. \ref{eq:hartlap}. In both cases we assume a Gaussian likelihood. We find a remarkably good agreement between both likelihoods, highlighting the negligible impact of the approximations involved in the NKA method on the final parameter estimates.} \label{fig:contours}
      \end{figure}
      Ultimately, the most important test to judge the accuracy of the
      analytical covariance matrix estimators implemented here is to study
      their impact on the posterior distribution of cosmological parameters
      derived from power spectrum measurements. Assuming flat priors \cgg{and
      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}}, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
      \begin{equation}\label{eq:parpos}
        -2\log p(\vec{\theta}|\hat{\bf C})=\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)^T\cdot{\rm Cov}^{-1}\cdot\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)+{\rm const.},
      \end{equation}
      where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.
      \cgg{Note that the low-$\ell$ non-gaussianities of real power spectra
        are not being taking into account with this approximation. They,
        however, have a lower statistical weight and will contribute less to
        the posterior distribution.}

@@

  \section{Discussion}\label{sec:discussion}
    Estimating accurate covariance matrices for projected two-point
    correlators is an ubiquitous problem in modern cosmology
    \cite{2013PhRvD..88f3537D,2013MNRAS.432.1928T,2017arXiv170609359K}, but it
    is particularly relevant for large-scale structure datasets. This problem
    is further complicated in this case, in comparison with e.g. CMB
    experiments, by 
    %two factors: 
    the fact that the fields involved are
    non-Gaussian at some level
    %and the higher complexity of the sky masks used in optical datasets.
    The main impact of survey geometry is the statistical
    coupling it induces between different Fourier/harmonic modes, which must
    be accurately characterized in order to obtain reliable estimates of the
    posterior distribution for cosmological parameters. In this paper we have
    focused on the impact of survey geometry on the dominant Gaussian (i.e.
    disconnected) part of the covariance matrix.

    We have described and generalized existing analytical approaches to
    estimate the covariance matrix for pseudo-$C_\ell$ power spectrum
    estimators
    \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and
    implemented them in the public code {\tt NaMaster}
    \cite{2019MNRAS.484.4127A}, making it straightforward to fully account for
    the effects of survey geometry on the data uncertainties. With these
    approximation\cgg{s}, computationally speaking, the problem of estimating a
    covariance matrix is as complex as that of computing the power spectrum
    itself, and scales with the number of pixels in the map as $N_{\rm
      pix}^{3/2}$. \cgg{We leave for future work the study of the impact of
      extra complications, as position-dependent noise, on their performance.}

    The main finding of this paper is the excellent performance of the
    analytical methods described in Section \ref{ssec:theory.pclcov}. We have
    shown that
    %, in spite of the higher complexity of large-scale structure masks,
    the NKA estimator (see Section \ref{ssec:theory.approx}) is able to
    recover the covariance matrix for all power spectra with non-zero signal
    expectation value (i.e. those involving the galaxy overdensity $\delta$ or
    the $E$-mode shear $\gamma_E$), as well as the posterior distribution for
    cosmological parameters, to a high degree of accuracy.

### 2019-09-27 draft/preprint_JCAP_v0.2/reply-to-comments-to-v0.1/reply-to-comments-to-v0.1.tex
implemented more changes from revision to v0.1

\title{Addressing the reviser concerns}
\author{Carlos Garc\'ia-Garc\'ia, Pilar Ru\'iz-Lapuente, David Alonso and\\ Miguel
  Zumalac\'arregui}

\begin{document}
\maketitle
We have carefully addressed the referee's questions and concerns. The
modifications we have performed, for each of the numbered requests, are as
follow:
\begin{enumerate}
  \item  Abstract. Line 9. Added explicit reference to CMB. We have changed
    \textit{expanding on existing algorithms} to \textit{expanding on existing algorithms \cgg{for CMB}}.
  \item  Page 8. After Eq. 2.44. Added GPE2 reference.
  \item  \dots
  \item  Last sentence of 3.3. Added \textit{or computed exactly as in [41,
      44]}. In the second to last paragraph of Section 4, we added at the end
    \textit{or exactly following Refs.~[41, 44]}.
  \item  In Section 3.3, before previous sentence, we changed \textit{If accurate
    covariances bars} to \textit{If accurate covariance matrix \cgg{elements}}.
\item  In Section 3.4, before Eq. 3.8, I have mentioned the Gaussian
  approximation of the likelihood: \textit{Assuming flat priors \cgg{and a
      fiducial Gaussian likelihood approximation [21]}}.
  \item  Second paragraph of Section 3.4. Changed \textit{for of two
      parameters} to \textit{for the \cgg{two} parameters}.
\end{enumerate}

Changes related to the CMB background:
\begin{enumerate}
  \item  
\end{enumerate}

Extra changes:
\begin{enumerate}
  \item  Changed $\Delta x^2$ to $\Delta^2 x$.
  \item  Introduction. Line 3. Updated the set of references for CMB.
  \item  Added HEALPix project url and credits in the Acknowledgments.
\end{enumerate}

>>>>

\title{Addressing the reviser concerns}
\author{Carlos Garc\'ia-Garc\'ia, David Alonso and Emilio Bellini}

\begin{document}
\maketitle
We have carefully addressed the referee's questions and concerns. The
modifications we have performed, for each of the numbered requests, are as
follow:
\begin{enumerate}
  \item  Abstract. Line 9. Added explicit reference to CMB. We have changed
    \textit{expanding on existing algorithms} to \textit{expanding on existing algorithms \cgg{for CMB}}.
  \item  Page 8. After Eq. 2.44. Added GPE2 reference.
  \item  Changed limits of Figure 6 and 7.
  \item  Last sentence of 3.3. Added \textit{or computed exactly as in [41,
      44]}. In the second to last paragraph of Section 4, we added at the end
    \textit{or exactly following Refs.~[41, 44]}.
  \item  In Section 3.3, before previous sentence, we changed \textit{If accurate
    covariances bars} to \textit{If accurate covariance matrix \cgg{elements}}.
\item  In Section 3.4, before Eq. 3.8, I have mentioned the Gaussian
  approximation of the likelihood: \textit{Assuming flat priors \cgg{and a
      fiducial Gaussian likelihood approximation [21]}}.
  \item  Second paragraph of Section 3.4. Changed \textit{for of two
      parameters} to \textit{for the \cgg{two} parameters}.
  \item  At the end of the first paragraph of Section 3.4,  after Eq. 3.8, I
    have added \textit{\cgg{Note that the low-$\ell$ non-gaussianities of real power spectra
        are not being taking into account with this approximation. They,
        however, have a lower statistical weight and will contribute less to
        the posterior distribution.}
}

\end{enumerate}

Changes related to the CMB background:
\begin{enumerate}
  \item  
  \item In Section 4, removed the statements about the higher complexity of LSS
    masks.
  \item  In the second paragraph of Section 4, we have added a comment on the
    effect of "real-wold" complications: \textit{With these
    approximation\cgg{s}, computationally speaking, the problem of estimating a
    covariance matrix is as complex as that of computing the power spectrum
    itself, and scales with the number of pixels in the map as $N_{\rm
      pix}^{3/2}$. \cgg{We leave for future work the study of the impact of
      extra complications, as position-dependent noise, on their performance.}
}
\end{enumerate}

Extra changes:
\begin{enumerate}
  \item  Changed $\Delta x^2$ to $\Delta^2 x$.
  \item  Before Eq. 2.12, added a comment on how the true power spectra must
    not change much over a bandpower. It says now \textit{Let us denote a given
          bandpower by its index $q$. \cgg{The number of bandpowers and their
            width are determined by the true underlying power spectra, which
            must not vary much within the bandpower}.}
  \item  Introduction. Line 3. Updated the set of references for CMB.
  \item  Added HEALPix project url, reference and credits in the
    Acknowledgments.
\end{enumerate}
